---
title: "R Notebook"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# Load Libraries and Data

```{r}
library(tidyverse)
library(naniar)
library(quanteda)
library(readxl)
library(ggthemes)
library(extrafont)
windowsFonts(sans = "Franklin Gothic Book")
loadfonts(device = "win")
loadfonts(device = "postscript")

setwd("X") 
ncap <- read_csv(
  file = "NCAPSurveySummary_SurveyResponsesDataset.csv", 
  col_names = T, 
  na = ""
) %>% 
  janitor::clean_names()

```


# Data Wrangling

```{r}
# Remove duplicates
ncap <- ncap %>% 
  distinct()
# Pivot Wide
ncap_wide <- ncap %>% 
  select(-respondent_count) %>% 
  pivot_wider(
    names_from = question, values_from = response
  ) %>% 
  janitor::clean_names()
# Remove NA's 
ncap_wide <- ncap_wide %>% 
  replace_with_na_all(condition = ~ .x %in% c("na", "NA", "N/A", "n/a", "Na", "N/a"))

# Clean up Colnames (dirty)
ncap_wide <- ncap_wide %>% 
  rename_with(
    ~ str_replace(.x, "(^[^_]*_[^_]*).*", "\\1"), 
    .cols = 25:ncol(ncap_wide)
  )
# Index a back-up vector to return to if needed 
back_up <- ncap_wide$wr28b_if
```

# Pre-Processing 
## Create Some pre-processing functions

### my_number_replacement

This function replaces digits in a string with their corresponding text.
It can be scaled out to include any range of numbers by adjusting the
the two associated indices: `numbers_text` & `digits`. The current
version captures the range of numbers 1:50.

```{r}
# Arguments 
## string ~ the string to be transformed 
## digits ~ the pattern of digits to be replaced
## number_text ~ the vector of replacement number texts 
### The final output will be the transformed string of text
my_number_replacement <- function(string, digits, number_text){
  reduce(
    seq_along(digits), function(string, i){
      str_replace_all(string, digits[i], number_text[i])
    }, 
    .init = string
  )
}
## Create some Indices for functions 
### Text for numbers
numbers_text <- c(
   "one", "two", "three", "four", "five", "six", "seven", "eight", "nine", "ten",
  "eleven", "twelve", "thirteen", "fourteen", "fifteen", "sixteen", "seventeen",
  "eighteen", "nineteen", "twenty", "twenty-one", "twenty-two", "twenty-three",
  "twenty-four", "twenty-five", "twenty-six", "twenty-seven", "twenty-eight",
  "twenty-nine", "thirty","thirty-one", "thirty-two", "thirty-three",
  "thirty-four", "thirty-five", "thirty-six", "thirty-seven", "thirty-eight",
  "thirty-nine", "forty", "forty-one", "forty-two", "forty-three", "forty-four",
  "forty-five", "forty-six", "forty-seven", "forty-eight", "forty-nine", "fifty", 
  "fifty-one", "fifty-two", "fifty-three", "fifty-four", "fifty-five", 
  "fifty-six", "fifty-seven", "fifty-eight", "fifty-nine", "sixty"
)
### Digits up to 50 
digits  <- paste0("\\b", 1:60, "\\b")
```


### my_ordinal_number_replacement
The function will replace all of the instances of ordinal numbers with their 
corresponding text (1st - first, 2nd - second, 3rd - third, etc.). Scaling this 
function out to capture a larger range of ordinals can be done as well, however 
please ensure you are checking to ensure the correct regex patterns are being 
implemented. The associated indices are: `ordinal_digits` & `ordinal_text`. 

```{r}
# Arguments
## string ~ the string to be transformed
## ordinal_digits ~ vector of the pattern of ordinal digits to be replaced
## ordinal_text ~ vector with the text of the replacement ordinal text
### The final output will be the transformed string of text
my_ordinal_number_replacement <- 
  function(string, ordinal_digits, ordinal_text){
  reduce(
    seq_along(ordinal_digits), function(string, i){
      str_replace_all(string, ordinal_digits[i], ordinal_text[i])
    }, 
    .init = string
  )
  }

# Create some Indices for the function 
## Ordinal text 
ordinal_text <- c(
  "first", "second", "third", "fourth", "fifth", "sixth", "seventh", "eighth", 
  "ninth", "tenth","eleventh", "twelfth", "thirteenth", "fourteenth",
  "fifteenth", "sixteenth", "seventeenth", "eighteenth", "nineteenth",
  "twentieth",  "twenty-first", "twenty-second", "twenty-third",
  "twenty-fourth", "twenty-fifth", "twenty-sixth", "twenty-seventh",
  "twenty-eighth", "twenty-ninth", "thirtieth","thirty-first", "thirty-second", 
  "thirty-third", "thirty-fourth", "thirty-fifth", "thirty-sixth",
  "thirty-seventh", "thirty-eighth", "thirty-ninth", "fortieth"
) 

## Ordinal Digits
ordinal_digits <- paste0(
  "\\b", 
  1:40, 
  str_sub(
    ordinal_text, 
    start = nchar(ordinal_text) - 1, 
    end = nchar(ordinal_text)
  ), 
  "\\b"
)
```


### my_stop_word_removal
This function will remove stopwords in a string. An index of stop words is 
created and can be adjusted to include or remove as needed.
```{r}
# Create a vector of stop words that to remove from the text
stop_words <- c("about", "above", "again", "against", "all", "am", "an", "and", 
                "any", "are", "aren't", "as", "at", "be", "because", "been", 
                "before", "being", "below", "between", "both", "but", "by", 
                "can't", "cannot", "could", "couldn't", "did", "didn't", "do", 
                "does", "doesn't", "doing", "don't", "down", "during", "each",
                "few", "for", "from", "further", "had", "hadn't", "has",
                "hasn't", "have", "haven't", "having", "he", "he'd", "he'll",
                "he's", "her", "here", "here's", "hers", "herself", "him",
                "himself", "his", "how", "how's", "i", "i'd", "i'll", "i'm",
                "i've", "if", "in", "into", "is", "isn't", "it", "it's", "its",
                "itself", "let's", "me", "more", "most", "mustn't", "my",
                "myself", "nor", "of", "other", "ought", "our",
                "ours", "ourselves", "out", "over", "own", "same", "shan't",
                "she", "she'd", "she'll", "she's", "should", "shouldn't", "so",
                "some", "such", "than", "that", "that's", "the", "their", 
                "theirs", "them", "themselves", "then", "there", "there's", 
                "these", "they", "they'd", "they'll", "they're", "they've", 
                "this", "those", "through", "to", "too", "under", "until", 
                "up", "very", "was", "wasn't", "we", "we'd", "we'll", "we're",
                "we've", "were", "weren't", "what", "what's", "when", "when's",
                "where", "where's", "which", "while", "who", "who's", "whom",
                "why", "why's", "will", "with", "won't", "would", "wouldn't", 
                "you", "you'd", "you'll", "you're", "you've", "your", "yours",
                "yourself", "yourselves", "work", "worked")

## Create Regex Pattern
stop_words <- paste0("\\b", stop_words, "\\b")

# Arguments 
## string ~ the string to be transformed 
## stop_words ~ a character vector of stop words to be removed from the string
### the final output will be the transformed string
my_remove_stopwords <- function(string, stop_words){
  reduce(
    seq_along(stop_words), function(string, i){
      str_remove_all(string, stop_words[i])
    }, 
    .init = string
  )
}

```

## Run Pre-Processing Steps 
```{r}
# Distinct Responses
ncap_wide %>% count(wr28b_if) 
```

```{r}
ncap_wide$wr28b_if <- back_up
```

```{r}
# Process Steps 
ncap_wide <- ncap_wide %>% 
  mutate(
    # Type cast to lowercase
    wr28b_if = tolower(wr28b_if), 
    # Punctutions
    ## Periods
    wr28b_if = str_replace_all(wr28b_if, "\\.", " "),
    ## Commas 
    wr28b_if = str_replace_all(wr28b_if, "\\,", " "),
    ## Apostrophes  
    wr28b_if = str_replace_all(wr28b_if, "\\'", " "),
    # Digits
    ## Capture one half 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b1/2\\b", 
      "one-half"
    ),
    ## Replace Ordinal Digits with number text first 
    wr28b_if = my_ordinal_number_replacement(
      wr28b_if,
      ordinal_digits, 
      ordinal_text
    ),
    ## Split digits from characters 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "(?<=\\D)(?=\\d)|(?<=\\d)(?=\\D)|(?<=\\d)(?=\\b)|(?<=\\b)(?=\\d)", 
      " "
    ),
    ## Replace all digits with their number text 
    wr28b_if = my_number_replacement(
      wr28b_if, 
      digits, 
      numbers_text
    ),
    # Remove Stop Words 
    #wr28b_if = my_remove_stopwords(
     # wr28b_if, 
      # stop_words
    #),
    # Standardize time periods 
    ## Hours 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(hours|hr|hrs)\\b", 
      "hour"
    ),
    ## Days
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(days)\\b", 
      "day"
    ),
    ## Weeks 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(wk|wks|weeks|weekly)\\b", 
      "week"
    ),
    ## Months
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(months|monthly)\\b", 
      "month"
    ),
    # Remove double spaces
    wr28b_if = str_replace_all(wr28b_if, "  ", " "), 
    # Remove Text NA's
    wr28b_if = case_when(
      str_detect(wr28b_if, "\\b(n/|n/a|na|not applicable)\\b") ~ NA, 
      str_detect(wr28b_if, "n/") ~ NA,
      str_detect(wr28b_if, "non-applicable") ~ NA,
      T ~ wr28b_if
    ),
    # Remove no or none text responses
    wr28b_if = case_when(
      str_detect(wr28b_if, "^\\b(no|none|nope)\\b") ~ NA, 
      str_detect(wr28b_if, 
                 "^\\b(not really|not sure|not regular|not that i am)\\b") ~ NA,
      T ~ wr28b_if
    ),
    # Remove the word yes
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(yeah|yes|yea|ya)\\b", 
      ""
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "^\\b(yeah|yes|yea|ya|)\\b", 
      ""
    ),
    # Remove Special Characters at the stary of a string
     wr28b_if = str_replace_all(
      wr28b_if, 
      "^[^A-Za-z0-9]+", 
      ""
    ),
    # Day Ranges 
    ## Monday - Friday 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "^\\b(m-f|mon-fri|mon - fri|m - f)\\b", 
      "monday through friday"
    ),
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(m-f|mon-fri|mon - fri|m - f)\\b", 
      "monday through friday"
    ),
    ## Monday - Saturday 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "^\\b(m-s|mon-sat|mon - sat|m - s)\\b", 
      "monday through saturday"
    ),
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(m-s|mon-sat|mon - sat|m - s)\\b", 
      "monday through saturday"
    ),
    # Days of the week
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(mon|m)\\b", 
      "monday"
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(tues|t)\\b", 
      "tuesday"
    ), 
        wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(wed|w)\\b", 
      "wednesday"
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(thurs|th)\\b", 
      "thursday"
    ),
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(fri|f)\\b", 
      "friday"
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(sat|s)\\b", 
      "saturday"
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\b(sun|s)\\b", 
      "sunday"
    ), 
     wr28b_if = str_replace_all(
      wr28b_if, 
      "^\\b(mon|m)\\b", 
      "monday"
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "^\\b(tues|t)\\b", 
      "tuesday"
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "^\\b(wed|w)\\b", 
      "wednesday"
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "^\\b(thurs|th)\\b", 
      "thursday"
    ),
    wr28b_if = str_replace_all(
      wr28b_if, 
      "^\\b(fri|f)\\b", 
      "friday"
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "^\\b(sat|s)\\b", 
      "saturday"
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "^\\b(sun|s)\\b", 
      "sunday"
    ),
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\bweekdays\\b", 
      "weekday"
    ), 
    wr28b_if = str_replace_all(
      wr28b_if, 
      "\\bweekends\\b", 
      "weekend"
    ),
    # Remove those with only digits in them 
    wr28b_if = case_when(
      !is.na(as.numeric(wr28b_if)) ~ "", 
      T ~ wr28b_if
    ),
    wr28b_if = str_trim(wr28b_if, side = "both"),
    wr28b_if = str_replace_all(wr28b_if, "\\s{2,}", " "), 
    wr28b_if = str_replace_all(wr28b_if, " ", "_"), 
    wr28b_if = str_replace_all(wr28b_if, "__", "_")
  )


```


## WR28B

```{r}
# Import Coded Responses to Match to Main DF 
wr28b <- bind_rows(
  read_xlsx("wr28b_if_open_text_brian.xlsx"), 
  read_xlsx("wr28b_if_open_text_dodds.xlsx")
)
```

```{r}
# Recode
wr28b <- wr28b %>% 
  mutate(
    rest_cycle_pattern = case_when(
      rest_cycle_pattern == 1 ~ "5 days on; 2 days off", 
      rest_cycle_pattern == 2 ~ "6 days on; 1 day off",
      rest_cycle_pattern == 3 ~ "14 days on; 1 day off", 
      rest_cycle_pattern == 4 ~ "Longer than 30 days with no days off", 
      rest_cycle_pattern == 5 ~ "Rest schedule established after three or more weeks", 
      rest_cycle_pattern == 6 ~ "Two or more cycles in same deployment",
      rest_cycle_pattern == 7 ~ "Other",
      T ~ NA
    ), 
    wr28b_if = str_replace_all(wr28b_if, "\\s{2,}", " "),
    wr28b_if = str_replace_all(wr28b_if, " ", "_"),
    wr28b_if = str_replace_all(wr28b_if, "__", "_"), 
    wr28b_if = str_trim(wr28b_if, "both")
  )
# Bring back to main DF
ncap_wide <- left_join(ncap_wide, wr28b %>% select(wr28b_if, rest_cycle_pattern), by = "wr28b_if")
```

# Create and export Data Viz for Slide Deck
# Lollipop
```{r}
ncap_wide %>% 
 count(rest_cycle_pattern) %>% 
  filter(!is.na(rest_cycle_pattern)) %>% 
    mutate(rest_cycle_pattern = case_when(
    rest_cycle_pattern == "Longer than 30 days with no days off" ~ 
      "30+ days with no day off", 
    rest_cycle_pattern == "Rest schedule established after three or more weeks" ~ 
      "Cycle established after 2+ weeks", 
    rest_cycle_pattern == "Two or more cycles in same deployment" ~
      "2+ cycles in same deployment", 
    T ~ rest_cycle_pattern
  )) %>% 
  ggplot(aes(x = fct_reorder(rest_cycle_pattern, n), y = n)) +
  geom_segment(
    aes(x = rest_cycle_pattern, xend = rest_cycle_pattern, y = 0, yend = n), 
    color = "#7aa5c1", linewidth = 1
  ) + 
  geom_point(color = "#005288", size = 10.5) + 
    geom_text(
      aes(label = n), size = 3, color = "#fcfdfd"
    ) + 
    xlab("") +
    ylab("") +
    ggtitle(label = "Rest Cycle Patterns Counts", subtitle = "WR28b")+ 
    coord_flip() +  
    theme(
    legend.position = "none",
    legend.title = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 11, color = "black"),
    axis.ticks.x =  element_blank(), 
    axis.text.x = element_blank(),
    plot.title = element_text(size = 14, face = "bold"),
    plot.background = element_blank(),
    # panel.grid = element_blank(),
    panel.background = element_blank(),
    axis.line.y =  element_line(colour = "black", linewidth = 1.5)
  ) 
ggsave("rest_day_lollipop.png")
```
# Barchart
```{r}
ncap_wide %>%
  count(rest_cycle_pattern) %>%
  filter(!is.na(rest_cycle_pattern)) %>%
  mutate(
    rest_cycle_pattern = case_when(
      rest_cycle_pattern == "Longer than 30 days with no days off" ~
        "30+ days with no day off",
      rest_cycle_pattern == "Rest schedule established after three or more weeks" ~
        "Cycle established after 2+ weeks",
      rest_cycle_pattern == "Two or more cycles in same deployment" ~
        "2+ cycles in same deployment",
      T ~ rest_cycle_pattern
    )
  ) %>%
  ggplot(aes(x = fct_reorder(rest_cycle_pattern, n), y = n)) +
  geom_bar(stat = "identity", fill = "#005288") +
  geom_text(aes(label = n), size = 4, color = "black", hjust = -.1, face = "bold") +
  xlab("") +
  ylab("") +
  ggtitle(label = "Rest Cycle Patterns Counts", subtitle = "WR28b") +
  ylim(c(0, 1800)) + 
  coord_flip() +
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 11, color = "black"),
    axis.ticks.x =  element_blank(), 
    axis.text.x = element_blank(),
    plot.title = element_text(size = 14, face = "bold"),
    plot.background = element_blank(),
    # panel.grid = element_blank(),
    panel.background = element_blank(),
    axis.line.y =  element_line(colour = "black", linewidth = 1.5)
  )
ggsave("rest_day_bar.png")
```

### Crosstab by Employeee Type
```{r}
ncap_wide %>%
  count(rest_cycle_pattern, dts_employee_type) %>%
  filter(!is.na(rest_cycle_pattern) & !is.na(dts_employee_type)) %>%
  mutate(
    rest_cycle_pattern = case_when(
      rest_cycle_pattern == "Longer than 30 days with no days off" ~
        "30+ days with no day off",
      rest_cycle_pattern == "Rest schedule established after three or more weeks" ~
        "Cycle established after 2+ weeks",
      rest_cycle_pattern == "Two or more cycles in same deployment" ~
        "2+ cycles in same deployment",
      T ~ rest_cycle_pattern
    )
  ) %>%
  ggplot(aes(x = fct_reorder(rest_cycle_pattern, n), y = n)) +
  geom_bar(stat = "identity", fill = "#005288") +
  geom_text(aes(label = n), size = 4, color = "black", hjust = -.1, face = "bold") +
  xlab("") +
  ylab("") +
  ggtitle(label = "Rest Cycle Patterns Counts", subtitle = "WR28b") +
  ylim(c(0, 1800)) + 
  coord_flip() +
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text.y = element_text(size = 11, color = "black"),
    axis.ticks.x =  element_blank(), 
    axis.text.x = element_blank(),
    plot.title = element_text(size = 14, face = "bold"),
    plot.background = element_blank(),
    # panel.grid = element_blank(),
    panel.background = element_blank(),
    axis.line.y =  element_line(colour = "black", linewidth = 1.5)
  ) + 
  facet_wrap(~dts_employee_type)
```

